---
title: "Class11"
author: "Michelle Janossy"
date: "5/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database 

The [PDB](http://www.rcsb.org/) is the main repository for biomolecular structure data. 

Here we examine the contents of the PDB:

>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
db <- read.csv("Data Export Summary.csv", row.names = 1)
head(db)
```

How many are X-Ray, etc...

```{r}
(db$Total/sum(db$Total))*100
```

What percent are Protein...

```{r}
sum(db$Proteins)/sum(db$Total)*100
```

We could also try the datapasta package and copy from website and "Addins" > "Paste as data.frame"...

```{r}
library(datapasta)

tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "Other", "NMR", "Multi Method",
                           "Electron Microscopy", "Total"),
              Proteins = c(126880, 256, 11062, 129, 2277, 140604),
         Nucleic.Acids = c(2012, 4, 1279, 5, 31, 3331),
    ProteinComplex = c(6547, 6, 259, 2, 800, 7614),
                 Other = c(8, 13, 8, 1, 0, 30),
                 Total = c(135447, 279, 12608, 137, 3108, 151579)
)


```

>Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1157 as of 2019-05-07 see:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=40D8F533

```{r}
library(bio3d)

pdb <- read.pdb("1hsg.pdb")
pdb
```

Atom selection is done via the function **atom.select()**

>Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues?

```{r}
prot.pdb <- atom.select(pdb, "protein", value = TRUE)
write.pdb(prot.pdb, file = "1hsg_protein.pdb")
lig.pdb <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(lig.pdb, file = "1hsg_protein.pdb")
```

Section 5

```{r}
aa <- get.seq("1ake_A")
```

```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

```{r}
head(hits$pdb.id)
```

```{r}
# Fetch PDBs
#files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
# Align structures
#pdbs <- pdbaln(files)
# Vector containing PDB codes
#ids <- basename.pdb(pdbs$id)
# Draw schematic alignment
#plot(pdbs, labels=ids)
```

