---
title: 'Class08: Machine Learning'
author: "Michelle Janossy"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's start with an example of running the **kmeans()** function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```
TO DO: 
-Use the kmeans() function setting k to 2 and nstart=20
-Inspect/print the results

> Q. How many points are in each cluster?
    30
    
```{r}
#km$size
```
    
> Q. What ‘component’ of your result object details
   - cluster size?
  - cluster assignment/membership?
  - cluster center?
  
```{r}
#km$center
```
  

> Plot x colored by the kmeans cluster assignment and
  add cluster centers as blue points

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

```{r}
plot(x, col=km$cluster)
points(km$centers, pch = 18, col = "blue", cex = 3)
```

## Hierarchical Clustering example

We must give the hclust() function a distance matrix not the raw data as an input

```{r}
#Distance matrix calculation
d <- dist(x)

#Clustering 
hc <- hclust(d)
plot(hc)
```

```{r}
plot(hc)
abline(h = 6, col = "red")
cutree(hc, k = 2)
cutree(hc, h = 6)
```

Another example but more real life like with overlapping points

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

> Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
> Q. How does this compare to your known 'col' groups?

```{r}
hc <- hclust(dist(x))
plot(hc)
abline(h=2, col="red")
abline(h=2.8, col="blue")
```

```{r}
gp2 <- cutree(hc, k=2)
gp3 <- cutree(hc, k=3)

gp2
gp3
```

```{r}
plot(x, col=gp3)
```

```{r}
table(gp2)
```

```{r}
table(gp3)
```

```{r}
table(gp2, gp3)
```

#PCA: Principal Component Analysis

We will use the **prcomp()** function for PCA

```{r}
##Example data
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata, 10) 
```

100 genes in this dataset
```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

Run our PCA analysis on the transpose 
```{r}
pca <- prcomp(t(mydata), scale = TRUE)
```


PCA plot
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2") 
```

Calculate the percent variance captures in each pC

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

head(pca.var.per)
```

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```

Example2 from class

```{r}
x <- read.csv("UK_foods.csv")

nrow(x)
ncol(x)
```

```{r}
head(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{r}
dim(x)
```

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))

mycols <- c("orange", "red", "blue", "darkgreen")

text(pca$x[,1], pca$x[,2], colnames(x), col = mycols)
```






























